---
title: "Consensus structure of the network communities"
author: ""
date: "`r format(Sys.time(), '%d %B %Y')`"
params:
  conf: "/data/cbprojectarchive/NGS_Riboswitch_655_Twister_Tet_screen13_RBB_2017/reports/report.conf"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    css: report.css
---

```{r setup, include=FALSE}
options(stringsAsFactors=FALSE)
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = FALSE
)
library(png)
source("report.helpers.R")
loadGlobalVars(params$conf)
```

### Load the results of the project

The DGE results and counts matrix are saved in `report.Rmd` in the file `r paste0(RESULTS, "/edger_objects.RData")`.

```{r LoadDEresults}
load(file=paste0(RESULTS, "/edger_objects.RData"))
```

### Read RNAfold data

Read the folding information from the ViennaRNA package:

```{r readRNAfold}
rnafold <- read.delim(paste0(RNAFOLD_OUT, "/rnafold.txt"), head=F)
rnafold$riboswitch <- gsub("^>(.+):([actgACTG]+)$", "\\1", rnafold$V1)
rnafold$switch     <- toupper(gsub("^>(.+):([actgACTG]+)$", "\\2", rnafold$V1))
rnafold$V4         <- gsub("[\\(\\)]", "", rnafold$V4)
#rnafold <- split(rnafold[, -1], rnafold$riboswitch)
```


### Network analysis {.tabset .tabset-fade .tabset-pills}

Redo the network analysis as in `report.Rmd`.

```{r network, results='asis'}
# retrieve the sequences for each pattern+contrast
s <- lapply(res, function(res.pattern) {
  lapply(res.pattern$LRT, function(res.pattern.contrast) {
    i <- res.pattern.contrast$table$FDR < .01
    data.frame(sequence=rownames(res.pattern.contrast$table)[i],
               logFC=res.pattern.contrast$table$logFC[i])
  })
})
s <- melt(s)
s$rnafold_key <- paste0(">", s$L1, ":", tolower(sub("_", "", s$sequence)))
s$full_sequence <- toupper(rnafold$V2[match(s$rnafold_key, rnafold$V1)])

# R> head(s)
#   sequence variable     value                          L2         L1
# 1  ACC_GGA    logFC 0.2739700 thiamine_100nM_vs_untreated TPP_TYPE_I
# 2  ACG_GGC    logFC 0.2161517 thiamine_100nM_vs_untreated TPP_TYPE_I
# 3  ACT_AGT    logFC 0.3819934 thiamine_100nM_vs_untreated TPP_TYPE_I
# 4  ATG_GGT    logFC 0.2323451 thiamine_100nM_vs_untreated TPP_TYPE_I
# 5  CAA_ACG    logFC 0.2935073 thiamine_100nM_vs_untreated TPP_TYPE_I
# 6  CCA_GGC    logFC 0.2187850 thiamine_100nM_vs_untreated TPP_TYPE_I
```

Calculate the network based on sequence distances and plot

```{r network2, results='asis'}
invisible({
  constructs <- split(s, s$L1)
  Map(constructs, names(constructs), f=function(x, construct.name) {
    cat("\n\n#### ", construct.name, " {.tabset}\n\n")
    treatments <- split(x, x$L2)
    Map(treatments, names(treatments), f=function(x, treatment.name) {
      cat("\n\n##### ", treatment.name, " {.tabset}\n\n")
      plot.net.edger.results(x$sequence, x$value, main="", sub="significant motifs at FDR 1%")

      if(length(x$sequence) < 2) {
        cat("less than 2 significantly enriched sequences")
        return()
      }

      # recalculate graph + communities as in plot.net.edger.results()
      d <- hamming(x$sequence)
      am <- ifelse(d == 1, 1, 0) # remove distant sequences and autoreferences
      if(all(am == 0)) {
        cat("there are no edges between vertices")
        return()
      }
      net <- graph_from_adjacency_matrix(am, mode="undirected")
      ceb <- cluster_edge_betweenness(net) 
      ceb.communities <- communities(ceb)

      # annotate the sequences with the community they belong to
      comm <- do.call(rbind, Map(ceb.communities, names(ceb.communities), f=function(x, x.name) data.frame(community=x.name, sequence=x)))
      x$community <- comm$community[match(x$sequence, comm$sequence)]

      # calculate avg FC of communities
      avg.fc <- tapply(x$value, x$comm, mean)

      # pick max and min communities 
      comm.max <- names(avg.fc)[which.max(avg.fc)]
      comm.min <- names(avg.fc)[which.min(avg.fc)]
      dir.create(paste0(RESULTS, "/network_community_folds/"), showWarnings=FALSE)
      file.suff <- paste0(RESULTS, "/network_community_folds/", construct.name, "_", treatment.name)
      write.csv(x[x$community == comm.max, ], paste0(file.suff, "_max.csv"))
      write.csv(x[x$community == comm.min, ], paste0(file.suff, "_min.csv"))

      # do MSA if 2+ seq (ClustalW), and predict secondary struct with RNA(ali)fold
      pred <- function(sequences, sequences.name, file.suff) {
        if(file.exists(paste0(file.suff, ".fa"))) {
          file.remove(paste0(file.suff, ".fa"))
        }
        invisible(
          Map(sequences, sequences.name, f=function(x, x.name) {
            cat(">", x.name, file=paste0(file.suff, ".fa"), sep="", fill=TRUE, append=TRUE)
            cat(x, file=paste0(file.suff, ".fa"), sep="", fill=TRUE, append=TRUE)
          })
        )

        if(length(sequences) > 1) {
          x <- system(paste0("~/bin/clustalw-2.1/clustalw2 -ALIGN -INFILE=", paste0(file.suff, ".fa")), intern=TRUE)
          x <- system(paste0("ml LocARNA ",
                             " && RNAalifold --color ", file.suff, ".aln", " > ", file.suff, ".mfe",
                             " && ps2pdf alirna.ps ", file.suff, ".pdf",
                             " && pdftopng ", file.suff, ".pdf ", file.suff,
                             " && rm alirna.ps"),
                      intern=TRUE)
        } else {
          x <- system(paste0("ml LocARNA ",
                             " && RNAfold -i ", file.suff, ".fa", " --id-prefix=rna.ps > ", file.suff, ".mfe",
                             " && ps2pdf rna.ps_0001_ss.ps ", file.suff, ".pdf",
                             " && pdftopng ", file.suff, ".pdf ", file.suff,
                             " && rm rna.ps_0001_ss.ps"),
                      intern=TRUE)
        }
        x <- system(paste0("pdftopng ", file.suff, ".pdf ", file.suff), intern=TRUE)
      }

      pred(x$full_sequence[x$community == comm.max], x$sequence[x$community == comm.max], paste0(file.suff, "_max"))
      pred(x$full_sequence[x$community == comm.min], x$sequence[x$community == comm.min], paste0(file.suff, "_min"))

      # read the resulting PNG and plot
      cols <- c("sequence", "variable", "value", "L2", "L1")
      cat("\n\n###### Max upregulated community (plot)\n\n")
      cat(paste0(file.suff, "_max-000001.png"), fill=TRUE)
      plot(0:1, 0:1, type="n", ann=FALSE, axes=FALSE)
      rasterImage(readPNG(paste0(file.suff, "_max-000001.png")), 0, 0, 1, 1)
      cat("\n\n###### Max upregulated community (table)\n\n")
      cat(knitr::kable(x[x$community == comm.max, cols], format="markdown"), sep="\n")
      cat("\n\n###### Min downregulated community (plot)\n\n")
      cat(paste0(file.suff, "_min-000001.png"), fill=TRUE)
      plot(0:1, 0:1, type="n", ann=FALSE, axes=FALSE)
      rasterImage(readPNG(paste0(file.suff, "_min-000001.png")), 0, 0, 1, 1)
      cat("\n\n###### Min downregulated community (table)\n\n")
      cat(knitr::kable(x[x$community == comm.min, cols], format="markdown"), sep="\n")
    })
  })
})
```

## Analysis Environment Log

* Analysis directory for current report: `r getwd()`
* Compilation time of knitr report: `r Sys.time()`

```{r sessionData, include=T}
sessionInfo()
```

